---
title: "RepData"
author: "Andriy Levytskyy"
date: "15/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Code for reading in the dataset and/or processing the data

```{r}
library(dplyr)

filename <- "RepData.zip"

# Checking if zip-file already exists.
if (!file.exists(filename)){
      fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
      download.file(fileURL, filename, method="curl")
}  

# Checking if un-zipped file exists
if (!file.exists("activity.csv")) { 
      unzip(filename) 
}

#Read data from activity.csv
activity<-read.csv("activity.csv")

activity$date<-as.Date(activity$date, "%Y-%m-%d")

head(activity)
```

## 2. Histogram of the total number of steps taken each day

```{r}
activity_by_date<-group_by(activity, date)

activity_sum_by_date<-summarise(activity_by_date, stepsSum=sum(steps))

hist(x=activity_sum_by_date$stepsSum)

```

## 3. Mean and median number of steps taken each day

```{r}
summary(activity_sum_by_date$stepsSum)

```

## 4. Time series plot of the average number of steps taken

```{r}
steps_by_interval <- aggregate(steps ~ interval, activity, mean)

plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
# find row with max of steps
max_steps_row <- which.max(steps_by_interval$steps)

# find interval with this max
steps_by_interval[max_steps_row, ]

```

## 6. Code to describe and show a strategy for imputing missing data

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs).
```{r}
sum(is.na(activity))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
activity_sum_na_by_date<-summarise(activity_by_date, stepsSum=sum(steps, na.rm = TRUE))

hist(x=activity_sum_na_by_date$stepsSum)

```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
activity_by_date$weekday <- weekdays(as.Date(activity_by_date$date))

activity_by_date$typeday <- "weekday"

activity_by_date$typeday[activity_by_date$weekday %in% c('Saturday','Sunday')] <- "weekend"

activity_by_date_week <- aggregate(steps ~ interval + typeday, data=activity_by_date, mean)

activity_by_date_weekday <- subset(activity_by_date_week, 
                                   activity_by_date_week$typeday=="weekday")
activity_by_date_weekend <- subset(activity_by_date_week, 
                                   activity_by_date_week$typeday=="weekend")

with(activity_by_date_weekday, plot(steps ~ interval, type = "l"))

with(activity_by_date_weekend, lines(steps ~ interval, type = "l", col = "red"))

legend("topright", col=c("black", "red"), lty = 1, lwd = 2, 
       legend = c("Weekday", "Weekend"))

```

## 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

